---
title: "Wine Quality Data Analysis"
author: "Sarah"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Wine Quality Distribution
The following code will evaluate the basic descriptive statistics of the Wine Quality variable of interest/ variable to be tested in the machine learning algorithms.

```{r}
#import and view dataset
library(readxl)
winequalityred <- read_excel("C:/Users/sarah/Desktop/info8000/winequalityred.xlsx", 
    col_types = c("numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))
QualityMax<-max(winequalityred$quality)
QualityMin<-min(winequalityred$quality)
QualityAverage<-mean(winequalityred$quality)
QualityMedian<-median(winequalityred$quality)
QualityMax
QualityMin
QualityAverage
QualityMedian
QualityMode <- names(which.max(table(winequalityred$quality)))
QualityMode
QualityStandardDeviation <- sd(winequalityred$quality)
QualityStandardDeviation
#as was stated in the UCI machine learning respository, there aren't many extreme values (very good/very bad wines). most of the wines are "Normal".

```

## Wine Quality Distribution Graph


```{r winequalityred}
ggplot(data=winequalityred, aes(x=quality)) +
    geom_histogram(binwidth=1, color='black', fill='white') +
    coord_cartesian(xlim=c(0,10)) +
    geom_vline(xintercept = QualityMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = QualityAverage, linetype=1, color='green', alpha=.5) +
    xlab("Wine Quality") +
    ylab("Number of Observations")
```

##Independent Variable Descripive Statistics
#Fixed Acidity
```{r winequalityred}
#fixedac
fixedacMax<-max(winequalityred$fixedacidity)
fixedacMin<-min(winequalityred$fixedacidity)
fixedacAverage<-mean(winequalityred$fixedacidity)
fixedacMedian<-median(winequalityred$fixedacidity)
fixedacMax
fixedacMin
fixedacAverage
fixedacMedian
fixedacMode <- names(which.max(table(winequalityred$fixedacidity)))
fixedacMode
fixedacStandardDeviation <- sd(winequalityred$fixedacidity)
fixedacStandardDeviation
```
#Volatile Acidity
```{r winequalityred}
volatileacMax<-max(winequalityred$volatileacidity)
volatileacMin<-min(winequalityred$volatileacidity)
volatileacAverage<-mean(winequalityred$volatileacidity)
volatileacMedian<-median(winequalityred$volatileacidity)
volatileacMax
volatileacMin
volatileacAverage
volatileacMedian
volatileacMode <- names(which.max(table(winequalityred$volatileacidity)))
volatileacMode
volatileacStandardDeviation <- sd(winequalityred$volatileacidity)
volatileacStandardDeviation
```
#Citric Acid
```{r winequalityred}
citricMax<-max(winequalityred$citricacid)
citricMin<-min(winequalityred$citricacid)
citricAverage<-mean(winequalityred$citricacid)
citricMedian<-median(winequalityred$citricacid)
citricMax
citricMin
citricAverage
citricMedian
citricMode <- names(which.max(table(winequalityred$citricacid)))
citricMode
citricStandardDeviation <- sd(winequalityred$citricacid)
citricStandardDeviation
```
#Residual Sugar
```{r winequalityred}
sugarMax<-max(winequalityred$residualsugar)
sugarMin<-min(winequalityred$residualsugar)
sugarAverage<-mean(winequalityred$residualsugar)
sugarMedian<-median(winequalityred$residualsugar)
sugarMax
sugarMin
sugarAverage
sugarMedian
sugarMode <- names(which.max(table(winequalityred$residualsugar)))
sugarMode
sugarStandardDeviation <- sd(winequalityred$residualsugar)
sugarStandardDeviation
```
#Chlorides
```{r winequalityred}
chloridesMax<-max(winequalityred$chlorides)
chloridesMin<-min(winequalityred$chlorides)
chloridesAverage<-mean(winequalityred$chlorides)
chloridesMedian<-median(winequalityred$chlorides)
chloridesMax
chloridesMin
chloridesAverage
chloridesMedian
chloridesMode <- names(which.max(table(winequalityred$chlorides)))
chloridesMode
chloridesStandardDeviation <- sd(winequalityred$chlorides)
chloridesStandardDeviation
```
#Free SO2
```{r winequalityred}
freesulfurdioxideMax<-max(winequalityred$freesulfurdioxide)
freesulfurdioxideMin<-min(winequalityred$freesulfurdioxide)
freesulfurdioxideAverage<-mean(winequalityred$freesulfurdioxide)
freesulfurdioxideMedian<-median(winequalityred$freesulfurdioxide)
freesulfurdioxideMax
freesulfurdioxideMin
freesulfurdioxideAverage
freesulfurdioxideMedian
freesulfurdioxideMode <- names(which.max(table(winequalityred$freesulfurdioxide)))
freesulfurdioxideMode
freesulfurdioxideStandardDeviation <- sd(winequalityred$freesulfurdioxide)
freesulfurdioxideStandardDeviation
```

#Total SO2
```{r winequalityred}
totalsulfurdioxideMax<-max(winequalityred$totalsulfurdioxide)
totalsulfurdioxideMin<-min(winequalityred$totalsulfurdioxide)
totalsulfurdioxideAverage<-mean(winequalityred$totalsulfurdioxide)
totalsulfurdioxideMedian<-median(winequalityred$totalsulfurdioxide)
totalsulfurdioxideMax
totalsulfurdioxideMin
totalsulfurdioxideAverage
totalsulfurdioxideMedian
totalsulfurdioxideMode <- names(which.max(table(winequalityred$totalsulfurdioxide)))
totalsulfurdioxideMode
totalsulfurdioxideStandardDeviation <- sd(winequalityred$totalsulfurdioxide)
totalsulfurdioxideStandardDeviation
```
#Density
```{r winequalityred}
densityMax<-max(winequalityred$density)
densityMin<-min(winequalityred$density)
densityAverage<-mean(winequalityred$density)
densityMedian<-median(winequalityred$density)
densityMax
densityMin
densityAverage
densityMedian
densityMode <- names(which.max(table(winequalityred$density)))
densityMode
densityStandardDeviation <- sd(winequalityred$density)
densityStandardDeviation
```
#PH
```{r winequalityred}
pHMax<-max(winequalityred$pH)
pHMin<-min(winequalityred$pH)
pHAverage<-mean(winequalityred$pH)
pHMedian<-median(winequalityred$pH)
pHMax
pHMin
pHAverage
pHMedian
pHMode <- names(which.max(table(winequalityred$pH)))
pHMode
pHStandardDeviation <- sd(winequalityred$pH)
pHStandardDeviation
```

#Sulfates
```{r winequalityred}
sulphatesMax<-max(winequalityred$sulphates)
sulphatesMin<-min(winequalityred$sulphates)
sulphatesAverage<-mean(winequalityred$sulphates)
sulphatesMedian<-median(winequalityred$sulphates)
sulphatesMax
sulphatesMin
sulphatesAverage
sulphatesMedian
sulphatesMode <- names(which.max(table(winequalityred$sulphates)))
sulphatesMode
sulphatesStandardDeviation <- sd(winequalityred$sulphates)
sulphatesStandardDeviation
```

#Alcohol
```{r winequalityred}
alcoholMax<-max(winequalityred$alcohol)
alcoholMin<-min(winequalityred$alcohol)
alcoholAverage<-mean(winequalityred$alcohol)
alcoholMedian<-median(winequalityred$alcohol)
alcoholMax
alcoholMin
alcoholAverage
alcoholMedian
alcoholMode <- names(which.max(table(winequalityred$alcohol)))
alcoholMode
alcoholStandardDeviation <- sd(winequalityred$alcohol)
alcoholStandardDeviation
```

##Variable Correlations
```{r winequalityred}

```

