---
title: "Wine Quality Data Analysis"
author: "Sarah"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Wine Quality
The following code will evaluate the basic descriptive statistics of the Wine Quality variable of interest/ variable to be tested in the machine learning algorithms.

```{r}
#import and view dataset
library(readxl)
winequalityred <- read_excel("C:/Users/sarah/Desktop/info8000/winequalityred.xlsx", 
    col_types = c("numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))
QualityMax<-max(winequalityred$quality)
QualityMin<-min(winequalityred$quality)
QualityAverage<-mean(winequalityred$quality)
QualityMedian<-median(winequalityred$quality)
QualityMax
QualityMin
QualityAverage
QualityMedian
QualityMode <- names(which.max(table(winequalityred$quality)))
QualityMode
QualityStandardDeviation <- sd(winequalityred$quality)
QualityStandardDeviation
#as was stated in the UCI machine learning respository, there aren't many extreme values (very good/very bad wines). most of the wines are "Normal".

```

##Wine Quality Distribution Graph


```{r}
library(ggplot2)
ggplot(data=winequalityred, aes(x=quality)) +
    geom_histogram(binwidth=1, color='black', fill='white') +
    coord_cartesian(xlim=c(0,10)) +
    geom_vline(xintercept = QualityMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = QualityAverage, linetype=1, color='green', alpha=.5) +
    xlab("Wine Quality") +
    ylab("Number of Observations")
```

##Independent Variable Descripive Statistics
#Fixed Acidity
```{r}
#fixedac
fixedacMax<-max(winequalityred$fixedacidity)
fixedacMin<-min(winequalityred$fixedacidity)
fixedacAverage<-mean(winequalityred$fixedacidity)
fixedacMedian<-median(winequalityred$fixedacidity)
fixedacMax
fixedacMin
fixedacAverage
fixedacMedian
fixedacMode <- names(which.max(table(winequalityred$fixedacidity)))
fixedacMode
fixedacStandardDeviation <- sd(winequalityred$fixedacidity)
fixedacStandardDeviation
```
#Fixed Acidity Distribution
```{r}
ggplot(data=winequalityred, aes(x=fixedacidity)) +
    geom_histogram(binwidth=1, color='black', fill='white') +
    coord_cartesian(xlim=c(0,16)) +
    geom_vline(xintercept = fixedacMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = fixedacAverage, linetype=1, color='red', alpha=.5) +
    xlab("Fixed Acidity") +
    ylab("Number of Observations")
```

#Volatile Acidity
```{r}
volatileacMax<-max(winequalityred$volatileacidity)
volatileacMin<-min(winequalityred$volatileacidity)
volatileacAverage<-mean(winequalityred$volatileacidity)
volatileacMedian<-median(winequalityred$volatileacidity)
volatileacMax
volatileacMin
volatileacAverage
volatileacMedian
volatileacMode <- names(which.max(table(winequalityred$volatileacidity)))
volatileacMode
volatileacStandardDeviation <- sd(winequalityred$volatileacidity)
volatileacStandardDeviation
```
##Volatile Acidity Distribution
```{r}
ggplot(data=winequalityred, aes(x=volatileacidity)) +
    geom_histogram(binwidth=.1, color='black', fill='white') +
    coord_cartesian(xlim=c(0,1.6)) +
    geom_vline(xintercept = volatileacMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = volatileacAverage, linetype=1, color='red', alpha=.5) +
    xlab("Volatile Acidity") +
    ylab("Number of Observations")
```

#Citric Acid
```{r}
citricMax<-max(winequalityred$citricacid)
citricMin<-min(winequalityred$citricacid)
citricAverage<-mean(winequalityred$citricacid)
citricMedian<-median(winequalityred$citricacid)
citricMax
citricMin
citricAverage
citricMedian
citricMode <- names(which.max(table(winequalityred$citricacid)))
citricMode
citricStandardDeviation <- sd(winequalityred$citricacid)
citricStandardDeviation
```
## Citric acid distribution
```{r}
ggplot(data=winequalityred, aes(x=citricacid)) +
    geom_histogram(binwidth=.1, color='black', fill='white') +
    coord_cartesian(xlim=c(0,1)) +
    geom_vline(xintercept = citricMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = citricAverage, linetype=1, color='red', alpha=.5) +
    xlab("Citric Acid") +
    ylab("Number of Observations")
```

#Residual Sugar
```{r}
sugarMax<-max(winequalityred$residualsugar)
sugarMin<-min(winequalityred$residualsugar)
sugarAverage<-mean(winequalityred$residualsugar)
sugarMedian<-median(winequalityred$residualsugar)
sugarMax
sugarMin
sugarAverage
sugarMedian
sugarMode <- names(which.max(table(winequalityred$residualsugar)))
sugarMode
sugarStandardDeviation <- sd(winequalityred$residualsugar)
sugarStandardDeviation
```
#sugar distribution
```{r}
ggplot(data=winequalityred, aes(x=residualsugar)) +
    geom_histogram(binwidth=.5, color='black', fill='white') +
    coord_cartesian(xlim=c(0,16)) +
    geom_vline(xintercept = sugarMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = sugarAverage, linetype=1, color='red', alpha=.5) +
    xlab("Residual Sugar") +
    ylab("Number of Observations")
```

#Chlorides
```{r}
chloridesMax<-max(winequalityred$chlorides)
chloridesMin<-min(winequalityred$chlorides)
chloridesAverage<-mean(winequalityred$chlorides)
chloridesMedian<-median(winequalityred$chlorides)
chloridesMax
chloridesMin
chloridesAverage
chloridesMedian
chloridesMode <- names(which.max(table(winequalityred$chlorides)))
chloridesMode
chloridesStandardDeviation <- sd(winequalityred$chlorides)
chloridesStandardDeviation
```
##Chlorides Distribution
```{r}
ggplot(data=winequalityred, aes(x=chlorides)) +
    geom_histogram(binwidth=.01, color='black', fill='white') +
    coord_cartesian(xlim=c(0,.7)) +
    geom_vline(xintercept = chloridesMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = chloridesAverage, linetype=1, color='red', alpha=.5) +
    xlab("Chlorides") +
    ylab("Number of Observations")
```

#Free SO2
```{r}
freesulfurdioxideMax<-max(winequalityred$freesulfurdioxide)
freesulfurdioxideMin<-min(winequalityred$freesulfurdioxide)
freesulfurdioxideAverage<-mean(winequalityred$freesulfurdioxide)
freesulfurdioxideMedian<-median(winequalityred$freesulfurdioxide)
freesulfurdioxideMax
freesulfurdioxideMin
freesulfurdioxideAverage
freesulfurdioxideMedian
freesulfurdioxideMode <- names(which.max(table(winequalityred$freesulfurdioxide)))
freesulfurdioxideMode
freesulfurdioxideStandardDeviation <- sd(winequalityred$freesulfurdioxide)
freesulfurdioxideStandardDeviation
```
##Free SO2 Distribution
```{r}
ggplot(data=winequalityred, aes(x=freesulfurdioxide)) +
    geom_histogram(binwidth=2, color='black', fill='white') +
    coord_cartesian(xlim=c(0,75)) +
    geom_vline(xintercept = freesulfurdioxideMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = freesulfurdioxideAverage, linetype=1, color='red', alpha=.5) +
    xlab("Free Sulfur Dioxide") +
    ylab("Number of Observations")
```

#Total SO2
```{r}
totalsulfurdioxideMax<-max(winequalityred$totalsulfurdioxide)
totalsulfurdioxideMin<-min(winequalityred$totalsulfurdioxide)
totalsulfurdioxideAverage<-mean(winequalityred$totalsulfurdioxide)
totalsulfurdioxideMedian<-median(winequalityred$totalsulfurdioxide)
totalsulfurdioxideMax
totalsulfurdioxideMin
totalsulfurdioxideAverage
totalsulfurdioxideMedian
totalsulfurdioxideMode <- names(which.max(table(winequalityred$totalsulfurdioxide)))
totalsulfurdioxideMode
totalsulfurdioxideStandardDeviation <- sd(winequalityred$totalsulfurdioxide)
totalsulfurdioxideStandardDeviation
```
##Total Sulfur Dioxide Distribution
```{r}
ggplot(data=winequalityred, aes(x=totalsulfurdioxide)) +
    geom_histogram(binwidth=10, color='black', fill='white') +
    coord_cartesian(xlim=c(0,300)) +
    geom_vline(xintercept = totalsulfurdioxideMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = totalsulfurdioxideAverage, linetype=1, color='red', alpha=.5) +
    xlab("Total Sulfur Dioxide") +
    ylab("Number of Observations")
```

#Density
```{r}
densityMax<-max(winequalityred$density)
densityMin<-min(winequalityred$density)
densityAverage<-mean(winequalityred$density)
densityMedian<-median(winequalityred$density)
densityMax
densityMin
densityAverage
densityMedian
densityMode <- names(which.max(table(winequalityred$density)))
densityMode
densityStandardDeviation <- sd(winequalityred$density)
densityStandardDeviation
```
##Density Distribution
```{r}
ggplot(data=winequalityred, aes(x=density)) +
    geom_histogram(binwidth=.001, color='black', fill='white') +
    coord_cartesian(xlim=c(.99,1.005)) +
    geom_vline(xintercept = densityMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = densityAverage, linetype=1, color='red', alpha=.5) +
    xlab("Density") +
    ylab("Number of Observations")
```

#PH
```{r}
pHMax<-max(winequalityred$pH)
pHMin<-min(winequalityred$pH)
pHAverage<-mean(winequalityred$pH)
pHMedian<-median(winequalityred$pH)
pHMax
pHMin
pHAverage
pHMedian
pHMode <- names(which.max(table(winequalityred$pH)))
pHMode
pHStandardDeviation <- sd(winequalityred$pH)
pHStandardDeviation
```
#pH Distribution
```{r}
ggplot(data=winequalityred, aes(x=pH)) +
    geom_histogram(binwidth=.1, color='black', fill='white') +
    coord_cartesian(xlim=c(2.5,4.5)) +
    geom_vline(xintercept = pHMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = pHAverage, linetype=1, color='red', alpha=.5) +
    xlab("pH") +
    ylab("Number of Observations")
```

#Sulfates
```{r}
sulphatesMax<-max(winequalityred$sulphates)
sulphatesMin<-min(winequalityred$sulphates)
sulphatesAverage<-mean(winequalityred$sulphates)
sulphatesMedian<-median(winequalityred$sulphates)
sulphatesMax
sulphatesMin
sulphatesAverage
sulphatesMedian
sulphatesMode <- names(which.max(table(winequalityred$sulphates)))
sulphatesMode
sulphatesStandardDeviation <- sd(winequalityred$sulphates)
sulphatesStandardDeviation
```
##Sulphates Distribution
```{r}
ggplot(data=winequalityred, aes(x=sulphates)) +
    geom_histogram(binwidth=.1, color='black', fill='white') +
    coord_cartesian(xlim=c(0,2)) +
    geom_vline(xintercept = sulphatesMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = sulphatesAverage, linetype=1, color='red', alpha=.5) +
    xlab("Sulphates") +
    ylab("Number of Observations")
```

#Alcohol
```{r}
alcoholMax<-max(winequalityred$alcohol)
alcoholMin<-min(winequalityred$alcohol)
alcoholAverage<-mean(winequalityred$alcohol)
alcoholMedian<-median(winequalityred$alcohol)
alcoholMax
alcoholMin
alcoholAverage
alcoholMedian
alcoholMode <- names(which.max(table(winequalityred$alcohol)))
alcoholMode
alcoholStandardDeviation <- sd(winequalityred$alcohol)
alcoholStandardDeviation
```
##Alcohol Distribution
```{r}
ggplot(data=winequalityred, aes(x=alcohol)) +
    geom_histogram(binwidth=.5, color='black', fill='white') +
    coord_cartesian(xlim=c(8,15)) +
    geom_vline(xintercept = alcoholMedian, linetype='longdash', alpha=.5) +
    geom_vline(xintercept = alcoholAverage, linetype=1, color='red', alpha=.5) +
    xlab("Alcohol") +
    ylab("Number of Observations")
```

##Variable Correlations
```{r}
wine.cor = cor(winequalityred)
library("Hmisc")
wine.rcorr = rcorr(as.matrix(winequalityred))
wine.coeff = wine.rcorr$r
wine.p = wine.rcorr$P
library(corrplot)
corrplot(wine.cor)
corrplot(wine.cor, tl.cex = 1, tl.col = "black", method = "number",
         addCoef.col = "black", number.digits = 1, number.cex = .5, 
         cl.pos = 'b', cl.cex = .5, addrect = .5, rect.lwd = 1,
col = colorRampPalette(c("red", "darkgrey","blue"))(100))
```
#correlation cont'd
```{r}
library(knitr)
kable(wine.cor, caption = "wine quality feature correlations")
```

